### 2015-10-15 reasoning/huiswerk/inl1019.tex
tried beginning with one but having troubles

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\author{Jappie Klooster}
\title{On proving things}
\maketitle

\section{Proof GKC rules}
\subsection{$L\to$}
To proof that if one of the premises of $L\to$ of $GKC$ is not true
(under all valuations) than the conclusion not true either.

The definition of $L\to$ in $GKC$:

\begin{scprooftree}{1}
	\AxiomC{$\Gamma \Rightarrow A, \Delta$}
	\AxiomC{$\Gamma, B \Rightarrow \Delta$}
	\BinaryInfC{$\Gamma, A \to B \Rightarrow \Delta$}
\end{scprooftree}

Let $I(x)$ be the interpretation function that determines if sequent x
is true.

Assume one of the premises is false while the conclusion remains true.
So $I(\Gamma \Rightarrow A, \Delta) = 0$
$I(\Gamma, A \to B \Rightarrow \Delta) = 1$

To get $I((\Gamma \Rightarrow A, \Delta) = 0$ if $\Gamma$ should have 
the value of $\neg A$
\subsection{$R\to$}
\end{document}