### 2015-10-28 agents/summery.tex
another sumarry for another part

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\lstset{language=Java}
\author{Jappie Klooster}
\title{Agents and you, a brief summary}
\maketitle

\section{Introduction}
This will be an introduction to agents coming from someone who followed
` hbo informatica' as pre study.

In here I will write down the things I think are important.

\section{Modal logic}

\section{APL}
Medhi dastani thought about the agent programming langauge (APL)

\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

\noindent
The rules can be:

\begin{itemize}
	\item Planning rules (goal $\to$ plan)
	\item Procedural rules (events $\to$ plan)
	\item Plan repair rules (plan $\to$ plan)
\end{itemize}

\subsection{Processing mental states}

\begin{itemize}
	\item Generate plans for received events.
	\item Generate plans for goals.
	\item Process exceptions and handle failures.
	\item Repair plans.
	\item Select plans for execution.
	\item Execute plans.
\end{itemize}

\section{Markov}

\end{document}