### 2015-10-28 agents/summery.tex
added configuration and operational semantics

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\begin{document}
\lstset{language=Java}
\author{Jappie Klooster}
\title{Agents and you, a brief summary}
\maketitle

\section{Introduction}
This will be an introduction to agents coming from someone who followed
` hbo informatica' as pre study.

In here I will write down the things I think are important.

\section{Modal logic}

\section{APL}
Medhi dastani thought about the agent programming langauge (APL)

\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

\noindent
The rules can be:

\begin{itemize}
	\item Planning rules (goal $\to$ plan)
	\item Procedural rules (events $\to$ plan)
	\item Plan repair rules (plan $\to$ plan)
\end{itemize}

\subsection{Processing mental states}

\begin{itemize}
	\item Generate plans for received events.
	\item Generate plans for goals.
	\item Process exceptions and handle failures.
	\item Repair plans.
	\item Select plans for execution.
	\item Execute plans.
\end{itemize}

\subsection{Operational semantics}
Operational semantics defines the computation steps a program
configuration may take. Operational semantics allows studying
of programming constructs in a rigorous manner. It facilitates proving
general properties about a language. It lies close to the implementation
of an interpreter. It facilitates model checking.

\subsubsection{Basic syntax}
The configuration $C$ evolves into configuration $C'$:

\[C \to C'\]

If premise $P$ holds transition $C \to C'$ can be derived:

\begin{prooftree}
	\AxiomC{$P$}
	\UnaryInfC{$C \to C'$}
\end{prooftree}

\subsection{Configuration}
configurations only mention things that change.
Multi-agent configuration where $A$ is an agent and $\chi$ is the
environment:

\[\langle \{A_1, \cdots, A_n\}, \chi\rangle \]

Individual agent configuration where $i$ is the agent number, $\sigma$ the
believes, $\gamma$ the goals, $\Pi$ the plans $\theta$ the assignments and
$\xi$ the events:

\[A_i = \langle i, \sigma_i, \gamma_i, \Pi_i, \theta_i, \xi_i \rangle \]

\section{Markov}

\end{document}