### 2015-11-02 agents/summery.tex
added a bunch of information, halfway the last slide of markov

\documentclass{article}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{amssymb}
\usetikzlibrary{automata,arrows,positioning,calc}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\newcommand{\brcell}[2][l]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}

\begin{document}
\lstset{language=Java}
\author{Jappie Klooster}
\title{Agents and you, a brief summary}
\maketitle

\section{Introduction}
This will be an introduction to agents coming from someone who followed
` hbo informatica' as pre study.

In here I will write down the things I think are important.

\section{Modal logic}

\section{APL}
Medhi dastani taught about the agent programming langauge (APL).
\subsection{Agent data structure}
This represents the mental state of a agent. \\
\begin{tabular}{ll}
	Name & Does what \\ \toprule
	Beliefs & Information available to the agent \\
	Goals & Objectives that the agent wants to achieve \\
	Events & Obsevations of (environmental) changes \\
	Capabilities & Actions that the agent can perform \\
	Plans & Procedures to achieve objectives \\
	Reasoning rules & Reason about goal and plans \\
\end{tabular} \\

\noindent
The rules can be:

\begin{itemize}
	\item Planning rules (goal $\to$ plan)
	\item Procedural rules (events $\to$ plan)
	\item Plan repair rules (plan $\to$ plan)
\end{itemize}

\subsection{Processing mental states}

\begin{itemize}
	\item Generate plans for received events.
	\item Generate plans for goals.
	\item Process exceptions and handle failures.
	\item Repair plans.
	\item Select plans for execution.
	\item Execute plans.
\end{itemize}

\subsection{Operational semantics}
Operational semantics defines the computation steps a program
configuration may take. Operational semantics allows studying
of programming constructs in a rigorous manner. It facilitates proving
general properties about a language. It lies close to the implementation
of an interpreter. It facilitates model checking.

\subsubsection{Basic syntax}
The configuration $C$ evolves into configuration $C'$:

\[C \to C'\]

If premise $P$ holds transition $C \to C'$ can be derived:

\begin{prooftree}
	\AxiomC{$P$}
	\UnaryInfC{$C \to C'$}
\end{prooftree}

\subsection{Configuration}
configurations only mention things that change.
Multi-agent configuration where $A$ is an agent and $\chi$ is the
environment:

\[\langle \{A_1, \cdots, A_n\}, \chi\rangle \]

Individual agent configuration where $i$ is the agent number, $\sigma$ the
believes, $\gamma$ the goals, $\Pi$ the plans $\theta$ the assignments and
$\xi$ the events:

\[A_i = \langle i, \sigma_i, \gamma_i, \Pi_i, \theta_i, \xi_i \rangle \]

\subsection{Transitions}
The confiugrations described \emph{what} can change where transition
will describe \emph{how} they can change.

$\vDash$ Describes a first order entailment relation. As I understand
this means $A \vDash B$ there exsits a $B$ in $A$ which has an entailment
to $A$. So $A$ is a set of things and a $B$ like thing exists in $A$.
\subsubsection{Belief update}
Execution of a belief update action $\alpha$ modifies the belief and goal
bases. $Update$ is an update function that is specified by the believe
update actions and $\gamma'=\gamma - \{\psi \in \gamma | \sigma' \vDash \psi\}$.
Which basically says that $\gamma'$ is equal to $\gamma$ minus the update.

\begin{prooftree}
	\AxiomC{$Update(\sigma, \alpha\theta)=\sigma'$}
	\UnaryInfC{$\langle i, \sigma, \gamma, \{(\alpha,id)\}, \theta, 
		\xi\rangle \to 
	\langle \langle i, \sigma', \gamma', \{\}, \theta, \xi\rangle$}

\end{prooftree}

So on the left side of the arrow, the plan is to do a believeUpdate $\alpha$
on the right side this is done so its removed from the plans and the believes
\emph{and} goals are modified. The goals are modified to because if a believe
with the same name as a goal becomes true than the goal is removed. (this is
a design decision btw, its just how 2apl works).

\subsubsection{Adopta goal}
Execution of a goal adopt action adopta(g) adds the goal to the beginning of
the goal base. If the goal $g$ trough assignment $\theta$ does not already
exists in goalbase.

\begin{prooftree}
	\AxiomC{$\sigma \nvDash g\theta$}
	\UnaryInfC{$\langle 
		i, \sigma, [\gamma_1, \cdots, y_n], \{(adopta(g),id)\}, \theta, \xi, 
	\rangle \to \langle 
		i, \sigma, [g\theta,\gamma_1, \cdots, y_n], \{\}, \theta, \xi, 
	\rangle$}
\end{prooftree}

So the agent comes accross a adopta method in the plan base $\Pi$, if
the associated believe of the goal is not true then you can start adding
it.
\subsubsection{Applying planning goal rules}
The structure of a PG-rule in 2apl is as follows:

\[\kappa <- \beta | \pi \]

Where $\kappa$ is the name of the plan.
$\beta$ is the condition that is applied before
executing the plan (basically a big if around the plan). $\pi$ is
the body of the plan, a set of instructions.

\begin{prooftree}
	\AxiomC{$\gamma \vDash \kappa \tau_1 \quad \& \quad
	\sigma \vDash \beta \tau_1 \tau_2$}
	\UnaryInfC{$\langle
		i, \sigma, \gamma, \Pi, \theta, \xi
	\rangle \to \langle
	i, \sigma, \gamma, \Pi \cup \{ \pi \tau_1 \tau_2 \}, \theta, \xi
	\rangle$}
\end{prooftree}

The transition can be executed if $\kappa$ is an active goal (this valuation
is stored in $\tau_1$ \emph{and} $\beta$ is true according to the believe 
base (this eveluation is stored in $\tau_2$). If all these things are true
then $\pi$ is added to the plan base $\Pi$ \emph{but} it will only be executed
if $\tau_1$ and $\tau_2$ return true.
\subsection{Exam}
some notes I gathered for possible exam questions:

What happens if $\gamma = pos(2,3)$ and $\sigma = pos(4,5)$, resolve with
state transitions.

\section{Markov}
Computerised decision makers must deal with an uncertain world.

\subsection{Expectation}
Excpectation is the average value you excpect to happen. For example if you
throw a dice the average value you should get is 3.5. Because:

\[1 * 1/6 + 2 * 1/6 + 3 * 1/6 + 4 * 1/6 + 5 * 1/6 + 6 * 1/6 = 3.5\]

You multiple all possible values time the chance of them happening and then
calculate the sum of them. Formally it is written like this:

\[ E[X]=Def \sum_{all x} xP\{X=x\} \]

So you can fill it in like:
\[E[X]= 1 * P\{X=1\} + \cdots + 6 *P\{X=6\}\]
\[E[X]= 1 * 1/6 + \cdots + 6 *1/6 = 3.5\]

\subsubsection{Properties of expectation}

\begin{itemize}
	\item $E[c] = c$
	\item $E[X+Y] = E[X] + E[Y]$
	\item $E[cX] = cE[X]$
	\item $E[XY] = E[X]E[Y]$ if and only if $X$ and $Y$ are independent.
	\item $E[X] \le E[Y]$, if $X \le Y$ almost surely.
\end{itemize}

\paragraph{Law of the onconcious statistician}
You have to prove this, its \emph{not} a property:

\[E[g(X)]=\sum_{all x} g(x)P{X=x}\]

where $g$ is any real-valued function. (It returns a value that is part
of the real set of numbers).

\subsection{Conditional expectation}
The expectation of $X | Y = y$ is called the \emph{conditional expectation}
of $X$ given $Y$, and is denoted by:

\[E[X|Y=y]\]

$E[X|Y=y]$ Is a number that depends on $y \in Y$. So you only get an
expectation $X$ if you have an y that satisfies the condition.
This yields a random variable $E[X|Y]$. Because of this it has all the
properties of a random variable.

\subsubsection{Probability mass functions}
The standart notation for probability mass functions is:

\[p_x(x): D(X) \to [0,1]: x \mapsto P\{X=x\} \]
\[p_y(y): D(Y) \to [0,1]: y \mapsto P\{Y=y\} \]
\[p_{xy}(x,y): D(X) \times D(Y) \to [0,1]: (x,y) \mapsto P\{X=x,Y=y\} \]
\[p_{x|y}(x|y): D(X) \times D(Y) \to [0,1]: (x,y) \mapsto P\{X=x|Y=y\} \]

For example:

\[ P\{X=x|Y=y\}=\frac{P\{X=x,Y=y\}}{P\{Y=y\}}\]

Is often written as:

\[P_{x|y}(x|y)=\frac{P_{xy}(x,y)}{P_y(y)}\]

\subsection{Total expectation}
The tower property, which is awesome:

Let $X$ and $Y$ be random variables then:

\[ E[E[X|Y]] = E[X] \]

The proof is in the slides.

\subsection{Markov to matrix}
A finite Markov process consisting of $n$ states can be represented by an 
$n\times n$ \emph{probablity transition matrix}.

Consider the following markov process:
\begin{center}
\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
\node[state]    (A)                     {A};
\node[state]    (B)[above right of=A]   {B};
\node[state]    (C)[below right of=A]   {C};
\path
(A) edge[loop left]     node{0.5}         (A)
    edge[bend left]     node{0.3}     (B)
    edge[bend right, left]     node{0.2}      (C)
(B) edge[bend left,right]    node{0.9}      (A)
    edge[bend left,right]    node{0.1}      (C)
(C) edge[bend right, right]    node{1}      (A);
\end{tikzpicture}

The probability matrix belonging to this process looks like the following:

\[
\begin{matrix}
		& A		& B		& C \\
	A	& 0.5	& 0.3	& 0.2 \\
	B	& 0.9	& 0		& 0.1 \\
	C	& 1		& 0		& 0 \\
\end{matrix}
\]

\end{center}

Adding the $A,B,C$ is optional, but I find it gives more overview. The first
cell reads, the probobality of going from a to a.

\subsubsection{Stochastic}
A matrix where every row sums to1 is called a \emph{stochastic matrix}. Every
transition matrix is a \emph{stochastic matrix}.

\subsection{Sample path}
A \emph{sample path} of a markov chain is a specific specific realistion of
that chain.

\[ X_0,X_1,X_2,\dots\]

The $X_i$ are random variables.

\subsection{Markov process}
A \emph{Markov process} is a stochastic process $X_0, X_1, X_2, \dots$ such 
that:

\[ P\{ X_{n+1} X_n, X_{n-1}, \dots\}=P\{X_{n+1}|X_n\}\]
for $n \ge 0$

\subsubsection{Remarks}
If states are numbers, then often state space is denoted by $I$, with 
typical elements $i,j,k$.

A markov process is sometimes called a markov chain. These concepts are
equal.

If $S$ is countably infinite, then the probability transition matrix
can be represented as an ``open ended'' stochastic matrix.
\[
	P =
\begin{pmatrix}
	p_{1,1}	& p_{2,1}	& p_{3,1} & \cdots\\
	p_{1,2}	& p_{2,2}	& p_{3,2}	& \cdots\\
	p_{1,3}	& p_{2,3}	& p_{3,3}	& \cdots\\
	\vdots	& \vdots	& \vdots		& \ddots \\
\end{pmatrix}
\]
With markov proceci it is extremly important to beclear about the
\emph{cardinality of the state space} (finite, discrete, continues) and
the \emph{nature of time} (discrete, continuous)

A \emph{dead state} or a \emph{dead end} is a state that does not have
any outgoing arcs. An \emph{absorbing  state} is a state that has precisely
one outgoing arc with probability 1 to itself.

\subsection{Chapman-Kolmogorov equations}

\[P^{(n)}_{ij} = \sum_{all\mbox{\ }k} P^{(n_1)}_{ik} P^{(n_2)}_{kj} \]

It is convenient to list the $n$-step transition probabilities in a matrix.
\[
	P^{(n)} =
\begin{pmatrix}
	p^{(n)}_{1,1}	& \cdots	& p^{(n)}_{n,3} 	& \\
	\vdots	& \ddots	& \vdots	& \\
	p^{(n)}_{1,n}	& \cdots	& p^{(n)}_{n,n}	& \\
\end{pmatrix}
\]

In particiular $P^{(1)} = P$.

\subsubsection{Lemma}
Suppose $k+l=n$ Where $k,l\ge1$. Then $P^{(k)} \cdot P^{(l)} = P^{(n)}$

(I think this was proven)

$P^{(n)} = P^n$, this was definitivley proven. Not sure why you would
prove removing brackets, but whatever.

\subsection{Classes}
No this aren't default private structures (its a c++ reference).

Classes make dealing with large markov chains easier by grouping nodes
together. \emph{Open classes} allow the process to escape the class once
it enters a \emph{closed class} it will remain there forever.
A class is \emph{Transient} if a process can never reenter it, a
state is recurrent if it can. A process is multichain if there are
multiple recurrent classes, if there is just one reccurent class
its uni-chain.

A critical transition is a transition between classes.

A closed class doesn't have to be recurrent if there are infinite cases.
Otherwise a closed class is always recurrent.
An openclass is always transient.

\subsection{Markov reward}
A \emph{Markov reward process} is a markov process with rewards on
transitions. these rewards are real valued.

Markov rewards processes are also known as \emph{markov reward chain} or
\emph{valued Markov chains}.

example

\[
	P =
\begin{pmatrix}
	0.2	& 0.8 & 0 	& \\
	0.3	& 0	& 0.7	& \\
	0	& 0 & 1	& \\
\end{pmatrix}
\]
With reward matrix:
\[
	R =
\begin{pmatrix}
	1	& 2 & 0 	& \\
	0	& 0	& 3	& \\
	0	& 0 & 0	& \\
\end{pmatrix}
\]

\subsubsection{Total reward}

Let $\omega$ be a realisation of a Markov process:

\[\omega = s_0, s_1, \dots\]

where $s_0$ is a start state.
The rewards that accumulate are $r_1, r_2, \dots$:

\[ s_0 \xrightarrow[\mbox{reward } r_1]{} s_1 \xrightarrow[\mbox{reward } r_2]{} s_2\]

\paragraph{Definition}

$r_k | (s_0=1) : $ reward at kth step, given $s_0 = i$.
$R|(s_0=i),$ the \emph{total reward} : $(r_1 + r_2 \dots)|s_0=i,$
$V(i),$ the \emph{value} of state i : $E[R|s_0=i]$.

\subsection{The bellman equation}

\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]

This system of linear equations enables us to compute state values in two
ways, by using algebra or by using iteration. Iteration you just wait until
convergence within a specified range. For this course iteration is enough.

\subsubsection{Utility pump}
A utility pump is a MRP in which at least one recurrent state possesses a
non-zero reward.
\subsubsection{Discount factor}
A \emph{dicsount factor} is a real number $0 \le \gamma < 1$. it represents
the probability to survive from one state to the next.

\paragraph{remarks} The discount factor is constant. The discount factor
does not impose a fixed cuttoff bound. It is a probablistic concept.

\subsubsection{Discounted reward}
Suppose a discount factor $0 \le \gamma < 1$. Then

\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + \gamma V(j)]\]

for every state $i$.

\paragraph{definition} A \emph{valuation} is a vector

\[v=Def
\begin{pmatrix}
	v(1) \\
	\vdots \\
	v(n)
\end{pmatrix}
\in \mathbb{R}^n
\]

\paragraph{definition} The \emph{expected immediate reward} at state $i$ is
the by $P_{ij}$ - probabilities weighted averaga of the $r_{ij}$ rewards:

\[C(i)=Def\sum_{all \mbox{\ }j} p_{ij} r_{ij} \]

The (vertical) vector of expected immediate rewards is denoted by $C$.

\paragraph{definition} The \emph{Bellman operator}:

\[B_p : \mathbb{R}^n \to \mathbb{R}^n : v \mapsto C + \gamma Pv \]

\subsection{Fixed point}
The value vector, $v^*$, of a discounted reward Markov process is a fixed
point of the $B_p$ operator:

\[ B_p(v)=v\]

\[
	v_0 \to B_p(v_0) \to B_p B_p (v_0) \to \dots \to B^n_P v_0
	\to \dots \to v^*
\]

\subsection{Example Markov reward process}
Consider the following markov reard process. Suppose a discount factor
$\gamma = 0.5$.

\[ \left\{ \begin{matrix}
		s: & s(r=1, p=0.4) & ; & a(r=4, p=0.6), \\
		a: & s(r=3, p=0.3) & ; & a(r=8, p=0.7), \\
		b: & s(r=2, p=0.5) & ; & b(r=5, p=0.5). \\
	\end{matrix}
\right. \]

\subsubsection{Transition diagram}
\begin{tikzpicture}[->, >=stealth', auto, semithick, node distance=3cm]
\tikzstyle{every state}=[fill=white,draw=black,thick,text=black,scale=1]
\node[state]    (S)                     {s};
\node[state]    (A)[right of=S]   {a};
\node[state]    (B)[left of=S]   {b};
\path
(S) edge[loop above]     	node{p=0.4, r=1}(S)
    edge[above]     		node{p=0.6, r=4}(A)
(A) edge[bend left,below]   node{p=0.3, r=3}(S)
	edge[loop above]     	node{p=0.7, r=8}(A)
(B) edge[above]    			node{p=0.5, r=2}(S)
	edge[loop above]     	node{p=0.5, r=5}(B);
\end{tikzpicture}
\subsubsection{Matrici}
\paragraph{Probability transition matrix}
\[
	P =
\begin{pmatrix}
	0.4	& 0.6 & 0 	& \\
	0.3	& 0.7 & 0	& \\
	0.5	& 0 & 0.5	& \\
\end{pmatrix}
\]
\paragraph{Immediate reward matrix}
\[
	R =
\begin{pmatrix}
	1	& 4 & 0 	& \\
	3	& 8	& 3	& \\
	2	& 0 & 5	& \\
\end{pmatrix}
\]
\subsubsection{Value evalutaion}
apply:
\[V(i)=\sum_{all\mbox{\ }j} p_{ij}[r_{ij} + V(j)]\]

\[ \left\{ \begin{matrix}
		V(s) = 0.4(1+\gamma V(s)) + 0.6(4+\gamma V(a)), \\
		V(a) = 0.3(3+\gamma V(s)) + 0.7(8+\gamma V(a)), \\
		V(b) = 0.5(2+\gamma V(s)) + 0.5(5+\gamma V(b)).
	\end{matrix}
\right. \]

With abusing notation:

\[ \left\{ \begin{matrix}
		s = 0.4(1+\gamma s) + 0.6(4+\gamma a), \\
		a = 0.3(3+\gamma s) + 0.7(8+\gamma a), \\
		b = 0.5(2+\gamma s) + 0.5(5+\gamma b).
	\end{matrix}
\right. \]

Just fill in the values recursevily to perform value iteration.

\subsection{Decision processes}
I'm gonna cover this subject poorly because its dead simple coming from an
IT background.

\begin{tabular}{ll}
	word & definition \\ \toprule
	decision tree & a simple directed tree stucture \\
	states & represented by verticies.\\
	Immediate rewards & represented by numbers along the edges. \\
	Decision alternatives & the outgoing edges, also called possible actions. \\
	Decision point & A state where a decision can be made. \\
	Finite decision tree & \brcell{A tree with a finite number of states, \\
		{\footnotesize Makes you wonder if there could be infinite number 
		of decisions in such a tree.}} \\
\end{tabular}
\\
The set of possible actions at state $i \in I$ is denoted by $A_i$.
The immediate reward for action $j$ at $i$ is denoted by $r_{ij}$.

\subsubsection{Policy}
A \emph{policy} specifies for every decision point $s$ a probability
distribution over $A_s$. A policy is typically denoted by $\pi, \tau\dots$.
If at $s$ the probability distribution puts all weight on one choice, the
policy is said to be \emph{pure} at $s$. Otherwise the policy is said to
be \emph{mixed} at $s$.

A \emph{run} of a policy is the \emph{actual execution} of that policy.
An \emph{realisation}, or \emph{sample path} of a policy is the trace of
that run.

The \emph{payoff} or \emph{total reward} of a run is the total sum of rewards
acuumulated in its realisation.

\subsubsection{Value function}
A \emph{value function} is any function $V: S \to \mathbb{R}$. The
\emph{expected payoff of a policy $\pi$}, given a certain starting state $s$:
	\[E[\pi|s_0=s]\]
	is called the value of $s$ under $\pi$, and is denoted by $V^\pi(s)$.
Value functions are compared in the following way:

\[ V_1 \le V_2 \Leftrightarrow_{Def} V_1(s) \le V_2(s), \mbox{for all states } s.\]

Policy $\pi$ is \emph{as good as} policy $\tau$ if $V^\tau \le V^\pi$. A policy
is \emph{optimal} if it is as good as any other policy. A value function is
\emph{optimal} if it is induced by a policy, and is as large as any other
such value function.

\subsubsection{Policy = strategy}
% this is the first what came to mind when I saw the policies.
In game theory, a policy is called a \emph{strategy} and a pure strategy is
sometimes called \emph{stationary} or \emph{deterministic}. A mixed strategy
is sometimes called \emph{probabilistic, non-stationary,} or
\emph{non-deterministic}.
% Imaginative with their non-*

An example of a pure policy is
\[ \tau = \{(r,b),(a,d),(b,e),(d,h),(f,i)\} \]

All the decisions are made, even if you already cut them off.
So you say (Start, End) for each node with choice.

Mixed policy:
\[ \tau = \{(r,\{(a, 0.2), (b,0.8)\}),(a,d),(b,\{(e, 0.7), (f,0.3)\}),
(d,h),(f,i)\} \]
% yes some really poor langauge design over here.

\subsubsection{Expected payoff}
$\pi$'s expected payoff, $E[\pi|s_0=r]$, can be computed recursivly with the
law of total expectation.

\subsubsection{Backward induction}
A decision between alternatives that leads to certain outcomes, is a certainty
itself.

Backward induction is applicable in \emph{directed acyclic graphs}(DAGs).

\subsubsection{Decision problem}
A \emph{decision problem} amounts to determining the set $\Pi^*$, where:

\[ \Pi = max_{\pi \in \Pi}(payoff(\pi))\]

Where $\Pi$ is the set of all policies.

For finite decision trees: Supppose at least one optimal policy exists.
If this policy is mixed, it may be replaced by an optimal pure policy,
such that, at every state, the payoff of the pure policy equals the expected
payoff of the mixed policy.

\subsubsection{Problems}
\paragraph{infinite width} no best policy, because you're valuating forever.
\paragraph{infinite depth} How to compare infinite total rewards (can't).
\paragraph{discount factor} The right decision is very dependent on the
discount factor. Counter intiuitive choices might be better for example.

\subsection{Utility pumps}
If there is no discount factor policy is irrelivant (the reward will be
infinite no matter what you do). A tool is needed to kill infinity.
For example a finiti horizon, discount factor, or long-run average
payoffs.

In this course we will just use discount factors.

\subsection{Exam}
 For a given decision problem with a certain
discount factor.A pure policy sufices.
Do policy iteration, (a single iteration), do value iteration
(single iteration).

\end{document}