### 2015-11-10 evolutionaryComputation/notes.tex
decided to throw all the homework just in one git repo,
Saves a lot of repo creation on github, and thus a lot of time

# seminars

talk about it half an hour
discuss 15 minutes after it.

Slides should be good enough.

First come first serve, pick top 3
In groups of ~3

# Practical assignments
Artifical benchmarks.
... next thursday, groups 1 or 2 students
Not very complicated, can use any programing langauge.

prefered to do it in groups.

Link, can download C code.

# What is it?
No domain knowlegde of your problem.
Black box algorithms,
	good thing, it works on a lot of problems
	bad thing its slow.

There exists hybred form which combines evolutionary
with specialised local algrotims.
	These are good at jumping out of local optimums

# Key concepts of darwanian system
Vaiation should exists in the copies, not to much not to little
Can add a little change or combine a solutions.

Inheritance, if the parents are good, then the childrens are to probably.

You have to destroy things. otherwise you never end up with something new.

Compitition, there should be competition otherwise there is no evolution.

# Evolutionary algoritm
1. Generate purly randomly or get a good start solution.
2. if you start with a 'good start' then you isolate your searchspace. if the
population stay similiar.

hybred no mutation, only selection compitition and inheritance.

while Not terminated do:
	1. select mates(T)
	2. x= generate offspring (T)
	3. evaluate population (x)
	4. T = select fittest.
return T

Don't select to *hard*, then after x generation your population will converge.

These algoritms can be very creative.

#Selection
##Tournamnet selection:
	S1 f1
	s2 f2 <- third
	s3 f3 <- best one
	s4 f4 <- second

	$p(t) -> p^s(t)$

	s3 f3
	s4 f3
	s3 f3
	s2 f2

The best will have 2, the wors 0, and the medium 1.
selection based on a grid, only compete against those next to you.

##Truncation selection:
Sort on fitness, take the top 1/x, and give them x copies for the new
population.

##Proportional select

$P^{select}(s_1) = frac{F_i}{\sum_{j=1}F_j}$

The next generation will be proportionally filled with how good the previous
was. This has the problem that the population will converge very fast.

Its not really for pragmatics, but better for theorticall analises.
So you should always use tournament or truncation.

You don't need to have all darwenin properties,
Inheritance, representation, compitition.

IRL function evaluation is usually expensive.

#Varation
The closer you can stay to your representation, then the crosover operators
preservers it. So binary representation is bad.

## MUtation
Just pick something and change it.

## crosover
The child inherints what the parents agree on,
where not the child can use random.
### uniform
flip a coin for every position

### 2point
pick 2 random points and cross them over.
positionally biased.