### 2015-11-10 inteligentagents/notes.tex

#Exam

You can use all material except that about eleborated excersizes.
You should be able to work with it. You can mark stuff.

#known
Stuff in mehthods of AI should be known
Modal logic is very important.

meyer & van der hoek, epistamic logic for AI and computer science, cambridge
univ press.
Very usefull for self study. Because it has excersizes with eleborations in
the back.

#Inteligent agents

Has to do with behaviour, try to mimic bahavior from the natural world in
software.

Agents, has many definitions but in this course its something with a
coginitive state.
Believe desires intentions - in a methorphorical stance, so not really

## Complete autonomy
Is kindoff useless, we need limited autonomy.
We need some kind of initiative or proactive or goal-directed.

##Mostly described in having mental states.
BDI 'strong notion of agency'

Show informational and motivational attitudes. an agent has a
motivation system of its own. It can be programmed, but then
the agent is on its own.

#Properties of agents.

##Situated
It can sense and reason about the environment.

They don't just give advice like exper systems

##reactive

## proactive

persuin goals, make plans.

## social
compete with eachorther.

#Objects vs agents

objects have some autonomy, but agents its more involved.

Agents are more 'subject' than 'object' in linguistic

objects can just be used, can be wrong but ok. An agent however
can only ask, so it can be refused. This isn't failing
its just the way it works.

... agents are objects with an attitude... wil of its own.
This is just an idea and there is an entire course about how we can do this.
It should be logical.

#Multi agent systems
After christmess.

There are quite difficult logics to describe their behavior.
You want multiple agents to wrok on a complex task, so you want to devide
the labour. if one fails others can take over.
Gracefull degradation. <- if somethign is wrong not everything is over

## why MAS

also paralism
roles of which a task can be done

## application

Softbots

### Robots
The rottordam harbour, the cranes are agent oriented. work in progress though.
Self driving cars, that could comminucate....

Replace the controller by an agent. Cognotive robotics, robots that think.
A robot steared by an agent.

Itel?? agents/robots exchanging goals and believes.

### Industrial
Airtraffic, one of the first applications of agents.
Manufactiring more agile, which can change easily

### Comericial
Automatic auctions. specifiy limits and the agent will negotiate.

### Entertainment
Virtual characters, more intelligent. Common sense.
Cinema <-- titanic was the first one that used agents.

# courses
##Philosophy
sloman: <- emotional robots

###Artitsotle
Practical sylogism, reasoning about actions.

### Denet
Strong AI, the computer can implement thinking like a human, as complex the
human brain.

###BRatman intetnions
Something to do with desires. But he thought about how to implement it with
software.

##Logical foundations
It needs to be combined. Quite complicated.

##Reasoning about action and frame
Frame problem, implementation of actions does not change much, but the rest does.

Situation calculus, <- to controll robots

##Agent architectures