### 2015-11-10 methods/reasoning/notes.tex

\documentclass{report}
\usepackage{booktabs}
\usepackage{graphicx}
\graphicspath{{images/}}

\usepackage{amsmath}

\begin{document}
\author{J. T. Klooster}
\title{Reasoning track notes}
\maketitle
\tableofcontents

\chapter{Examples}

\begin{itemize}
	\item Software for medical diagonostics
			Figure out how and what diseases a patiant has.
	\item Software for games
			This is different from diagnostics.
\end{itemize}

there is often a common core. After this course it should become easier to
work with logic because you've developed your own as a refernce pont

\chapter{Logic}
stuff you need to know

\begin{itemize}
		\item Classical logic, important
		\item Epistemic logics
		\item modal logics, not really that important
		\item logics for linguistics
		\item fuzzy logics, seems important?
\end{itemize}

Logic should be decidable, can you solve the problem, for example:
if you have $L$, does $A$ hold in $L$?
does $L$ hold in $A$ \\

It might be difficult to combine this with expressive power.

\section{Predicate logic}

Predicate logic is undeidable. Propositional logic is but its weaker. \\

Description logic, is between predicate and propositional. But it isn't
used in this course. Description logic has been devloped to be used in practice,
meaning it wasn't really developed my mathematicans. But its very succesfull but.

\subsection{proofs}

If it is asked that something is proofed take effort to do it. No sloppy work!

\subsection{propoistional logic}

\begin{tabular}{@{}lll@{}}
$\Rightarrow$ & implication & if A is true, then B is true \\
$\iff$ & bi-implication

\end{tabular}

\subsubsection{CPC}
Clasical propositional logic $\wedge$ clasical propositional computation

\section{decidable}

Logic is decidable if there is a turing machine when presented with a formula
$\phi$ as input, outpus 1 if $\phi$ holds in the logic and 0 otherwise. \\

Valdity in classical propoistional logic CPC is decidable. \\
Validity in claical logic CPQ is undicidable \\

Something with P vs NP

\section{sound and complete}
A proof system is sound and complete for a sementics if and only if formula
is true acording to the semantics exeactly when it is dervable in the proof
system.

\section{Hilbert}

hilbert is difficult to work with. Despair in 5. 4. 3. \dots

There are a bunch of axioms which just define the operators I think.
And there is one rule the Modus Ponens.

\subsection{axioms}

all the actioms have a name.

W says weakining and E says Exchange.

\subsection{nasty part}

finding the right axiom

\subsection{premises}
also called assumption.

\subsection{Lemma}

\section{Proof}
A proof of $A$ is $A_1 \dots A_n$ such that:
\begin{itemize}
	\item $A_n=A$
	\item For all $1<= i <= n$ $A$ is an axiam in system \\
		there are $j,h<i$ such that $A_j A_h / A_i$
\end{itemize}

\end{document}