### 2015-11-12 evolutionaryComputation/notes2.tex
notes from the class

#dominio convergence...

linear function = no interaction between the bits.

functions with nonlinear interaction are problematic, because
you can't do bitlfips to see if the situation improves.

If a GA wants to be better than a bit flipper it has to be able to
interact with higher order blocks.

# Schemata
is a  Similirarity subset

Look at schemata to see how a GA works

$m(h, t)$, m counts the number of instances that start with h, t is the
generations.

## schema growth by selection
### Porpotionselect
DOn't use proportion select!

### tournement select
don't have nice forumula

### Schema distruption by mutation
Mutation is destructive on schemata
h=1xx0xx1xxx
with probability P I'm gonna flip one of h's bits

the survival probaility of a schemata is the probaility of not flipping
those bits

A schema with a higher order can be distrupted more easily

#### 1 point crosover
worst case definition.

#### uniform crosover
Again worst case, assume if its touched its destroyed.
so either touch them all or don't touch them all.

order should be small to survive in the schema

## Schema theorom
m(h,t+1), the next generation

if the net growth factor is bigger then 1, it'll grow
selection = important for growth factor
destruction is also important, ie mutation

# Holland
proposed genetical algoritms.
Schematas are S curves, after a while the strongest 'genes'/bits will converge
and compete with each other, so you'll have a domino effect.

## Bandit problem
two normal distribution in which you can sample but you want to reduce
the sample size of the wrong distribution.

GA will figure out the best normal, but it doesn't maximise the sum
so there is some debate going on about holland's method.

# Building blocks
any search algorithm has a bias.
small population size with high selection presure won't work.

in a GA you don't want to rely on mutation
# loose & thight
thight things that are close are importent ie (bit form an integer)
loose the order of the bits doesn't matter (I imagine some sort of bit mask)

# deceptive vs non deciptive
has to do with the schema's. If you look of the schema averages of all the
schama's of the blocks. and ignore the optimum, you'll drift to the wrong
side, with non-deceptive trap functions you won't.

with non-deceptive you'll drift towards the optimum instead of the local
optimum.

# practical

## Why do we do this?

Understand about what things we're talking about. get an understanding of
the basic concepts.

## assignments

### uniformly scaled counting ones function
Count the number of ones. Its easy to model
### linearly scaled counting ones function
first has fitness of 1, the second of 2 times 1 etc...

### trap functions
it uses blocks to get trapped in a local optimum.

#### Deceptive
it will probably get stuck in a local optimum

#### non deciptive
should drift towards the optimum

### randomly linked
randomly linked that the bits are placed at random position on the binary
string.

random, at the beginning. Not for the contest, but just for the creation of
new ones?
two possible methods:
	* For the creation of all subject create 1 global random mapping
	* For each instance create his own random mapping, this has the
	problem however of deciding which random mapping to use for the child:
		Father, mother or another new random.
This wasn't discussed in class SO I asume a global mapping has to be used.
But here I assume so maybe an email to the teacher could help, or just do
both methods and discuss them in the report.

### selection
You can play with it, crank it up untill you cancel all the deciptiveness is
fine.