### 2015-11-16 inteligentagents/notes3.tex
notes of intelligent agents lecture 3

\documentclass{article}
\begin{document}
\author{Jappie Klooster}
\title{Cohen \& Levesque}
\maketitle
\section{Langauge}
DONE = just done
; = sequence
? = test if its true (groovy truth)
+ = is a non detrmenistic choice, collect them in an autcome set.
* = is done a finite but arbitrary number of times, including 0

if $\phi$ then $\alpha_1$ else
$\alpha_2$ = ($\phi? ; \alpha_1) + (\neg \phi?;\alpha)$

while $\phi$ do $\alpha$ = $(\phi?;\alpha)^*;\neg\phi?$

Pz = z has property P, a predicate attom

\subsection{Semantics}
$\omega$ = variables assginemnts
$P$ = is the set of omega $\in \omega$
$E$ = is a set of primitive events
$Agt = Agt \in [E \to p]$ specifies the agent of an event $[E \to P]$ is a
function that takes an E that gives a P.
$T \subset [Z \to E];$ a set of possible worlds, give a time number and get

i = agent
n = time

\subsubsection{Predicate}
c constants
f functions
Q predicate smbols

$\phi$ interperts predicate symbols

\[\phi(c) = d \in \Theta\]
\[\phi(f) = f D_1x\dots D_l \to D_h\]
\[\phi(Q^{n,x,ne}) = R \subseteq D_1x\dots D_l \]

\subsubsection{satisfaction}

\[M,\sigma,v,n[\alpha] m\]
In the model `$M,\sigma,v,n$', $\alpha$ lies betweeen n and m.

B = `normal` modal accesibility relation indexed by the time $n$
and the agent $v(i)$. they write it like:

\[<\sigma, n>B[v(i)] \sigma\]

Better would be:
\[\sigma B_{v(i)}^n \sigma\]

\subsection{Constraints}
\subsubsection{Believes/Goals}
KD45.

\[G\subseteq B\]
means $\vdash {Bel}_i \phi \to {Goal}_i\phi$, but not the otherway around

\subsection{Achievement goal}
Doesn't believe $\phi$ is true yet, but believes it could be true later.
There exists some action to make $\phi$ true.

\subsubsection{Asumption}
Goals will end some time in the future

\subsection{Persistent goals}
Drop the goal if its true, or its impossible.

Competence means that if he believes it then its true.

\section{Rao Georgeff}
\subsection{Time logics}
Brancing vs linear, points vs intervals,
discrete (real numbers) vs continues (complex number)
We'll only look into the future.
\end{document}