### 2015-11-23 inteligentagents/notes5.tex
added notes of Intelligent Agents

\documentclass{article}
\begin{document}
\author{Jappie Klooster}
\title{KARO logic}
\maketitle

There are indeed unkown unkowns. But not if you have
$\neg K \phi \to K \neg K \phi$

\section{the names are just the axioms}
jjK = ??
D = $\neg B \bot$
4 = $B\phi \to BB \phi$
5 = $\neg B \phi \to B\neg B \phi$

\section{Dynamic logics}
these actions can be non deterministic\ldots

Accessibility relation

$\alpha;\beta$ = sequentu\alphal compisition
$\alpha+\beta$ = nondeterministic choice
$\alpha*$ = execute \alphar\betaitr\alphary num\betaer of times

s' = a different variable to s, so instead of s' you can use [a-z].

Dynamic logic is complete!

does $[\alpha] ( \phi \to \varphi) \to ([\alpha] \phi \to [\alpha] \varphi)$
mean I can subtitute  $[\alpha] ( \phi \to \varphi)$ with
$([\alpha] \phi \to [\alpha] \varphi)$

$[\alpha*]\phi\to\phi$ do nothing its true
$[\alpha*]\phi\to[\alpha][\alpha*]\phi$ do once and it becomes true
\subsection{with induction}
$[\alpha*](\phi \to [\alpha]\phi)\to(\phi\to[\alpha*]\phi)$ 
$\phi\to[\alpha]\phi$
$[\alpha](\phi\to[\alpha]\phi)$
$[\alpha, \alpha](\phi\to[\alpha]\phi)$
$\vdots$

\section{Karo}
Opertunity vs capability: I have the opertunity to climb the dom tower\ldots

\subsection{KBD}
Knowledge Believe Desires.

$R_K = S5$
$R_B = KD45$
$R_B \subset \R_K = \vDash K\phi \to B\phi$
$R_D =$ no constraints, so sytem K.

infinite computation = $\nill$???

$R_\alpha(\nill)=\nill$ {divergence failure, or abort}

A = Agent is able to
$A\alpha \LeftRightarrow C(M,w)$

$<\phi?>\psi \LeftRightarrow (\psi \wedge \phi)$, if there is an accassible world
were $\psi$ holds and $\phi$ can be tested then in the world $\phi$ and $\psi$ hold.

Holds in general:

\[[\alpha]\phi \wedge <\alpha>\top\to <\alpha>\phi\]

If $\alpha$ is deterministic.

\[ <\alpha>\phi \to [\alpha]\phi \wedge <\alpha>\top\]

\end{document}