### 2015-12-06 evolutionaryComputation/notes4.tex
notes of lesson 5

\documentclass{article}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage{graphicx}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{tikz}
\usepackage{listings}
\usepackage{hyperref}
\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\newcommand{\brcell}[2][l]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\begin{document}
\author{Jappie Klooster}
\title{Evolutionary stretegies}
\maketitle

\section{difference from GA}
GA's are local searches, while ES are global searches.
\section{Comenly used settings}
Coveriance = the angle, no coveriance means semetrical

The closer to the optimum the smaller your variance should be. This reduces
the chance of jumping over the optimum.

Normally distributed = gausian

\section{mutation}

\section{reccombination}
the parents is the entire population

\section{selection}
 methods $(u,\lambda)$ and $(u+\lambda)$, $(u+\lambda)$ appears better from
 an optimization standpoint but self adoptation works worse with this one.

feet4, each edge = a weight, every neuron is a function.

\section{Combinatorial}
is a finite set of solutions and you have to optimize it.

\subsection{Local search}
always try to improve the current solution by applying small changes.

Meta heuristics, for example taboo list

\section{Second practical}

MLS, ILS and GLS for graph Bipartioning.

the easiest thing to do is how long and see on waht
percentage they from the total??

Graph biportitioning, cut the graph into 2 groups
of 250

if the hamming distance is bigger than 2, just take
the complement of the 2?

\section{Knapsack}
Adaptive pursuit, pick one random with probability x, change
probability while running.

\section{Adaptive pursuit}
probability matching is used most often in literature, but its not
really that good.

Markov!

\subsection{Strategy}
the learning pase is a prarameter you have to play with.

\section{Time complexity}
Convergence speed: easy problems easier to calculate

population should be large enough, reasonably.

Gentic programming uses extreme high selection pressure
	creating lisp with GA's.

Using this allows much better scalimg of the algortihm.
\end{document}