### 2016-01-23 evolutionaryComputation/summery.tex
added about ada pursuit and some about modaling the behavior

\documentclass[10pt,landscape]{article}
\usepackage{multicol}
\usepackage{calc}
\usepackage{ifthen}
\usepackage[landscape]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb}
\usepackage{color,graphicx,overpic}
\usepackage{hyperref}

\pdfinfo{
  /Title (example.pdf)
  /Creator (TeX)
  /Producer (pdfTeX 1.40.0)
  /Author (Seamus)
  /Subject (Example)
  /Keywords (pdflatex, latex,pdftex,tex)}

% This sets page margins to .5 inch if using letter paper, and to 1cm
% if using A4 paper. (This probably isn't strictly necessary.)
% If using another size paper, use default 1cm margins.
\ifthenelse{\lengthtest { \paperwidth = 11in}}
    { \geometry{top=.5in,left=.5in,right=.5in,bottom=.5in} }
    {\ifthenelse{ \lengthtest{ \paperwidth = 297mm}}
        {\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
        {\geometry{top=1cm,left=1cm,right=1cm,bottom=1cm} }
    }

% Turn off header and footer
\pagestyle{empty}

% Redefine section commands to use less space
\makeatletter
\renewcommand{\section}{\@startsection{section}{1}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%x
                                {\normalfont\large\bfseries}}
\renewcommand{\subsection}{\@startsection{subsection}{2}{0mm}%
                                {-1explus -.5ex minus -.2ex}%
                                {0.5ex plus .2ex}%
                                {\normalfont\normalsize\bfseries}}
\renewcommand{\subsubsection}{\@startsection{subsubsection}{3}{0mm}%
                                {-1ex plus -.5ex minus -.2ex}%
                                {1ex plus .2ex}%
                                {\normalfont\small\bfseries}}
\makeatother

% Define BibTeX command
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

% Don't print section numbers
\setcounter{secnumdepth}{0}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0pt plus 0.5ex}

%My Environments
\newtheorem{example}[section]{Example}
% -----------------------------------------------------------------------

\begin{document}
\raggedright
\footnotesize
\begin{multicols}{3}

% multicol parameters
% These lengths are set only within the two main columns
%\setlength{\columnseprule}{0.25pt}
\setlength{\premulticols}{1pt}
\setlength{\postmulticols}{1pt}
\setlength{\multicolsep}{1pt}
\setlength{\columnsep}{2pt}

\begin{center}
     \Large{\underline{Evolutionary Computing}} \\
\end{center}

\section{Genetic algorithms}
%TODO: crossover methods I don't understand
%TODO: schemas, buidling block theory
\subsection{Schema}
$o(h)$: schema order $o(11\#\#0)$ ie nonwildcard \# elements.
$\delta(h)$: schema defining length, maximum distance between two defining
symbols. $m(h,t)$: number of schema h instances at generation t.
$f(h,t)=\sum_{i\in P}f_i$ schema fitness is average fitness of members.
\subsubsection{Growth by selection}
reproduction ratio $\phi(h,t)=\frac{m(h,t^s)}{m(h,t)}$.
\paragraph{Proportionate selection}
$f_i$ fitness ind i. Probability individiual i selected: $\frac{f_i}{\sum f_i}$.
$N$ pop size. Expected number of copies
$\frac{f_i}{\sum f_i} \cdot N = \frac{f_i}{f(t)}$.
Expected number of copies of schema $h$ members:
$m(h,t^s) = m(h,t)\phi(h,t)=m(h,t)\frac{f(h,t)}{f(t)}$.
\paragraph{Tournament selction}
tournament size $s: 0 \leq \phi(h,t) \leq s$.

\subsubsection{Distruption by mutation}
probability bit flipped: $p_m$.
schema $h$ survives iff all the bit values are \emph{not} mutated
$p_{survival}=(1-P_m)^{o(h)}$. for small values $p_m << 1$,
$(1-p_m)^{o(h)}\approx 1 - o(h)\cdot p_m$.
disruption factor $\epsilon(h,t)$ by mutation: $\epsilon(h,t)=o(h)\cdot P_m$.

\subsubsection{Distruption by recombination}
Probablity crossover applied $P_c$. Bit swap probability: $p_x$.

\emph{1-point crossover} schema $h$ survives iff cutpoint \emph{not}
withing defining length $\delta$: $p_{survival} = 1-\frac{\delta(h,t)}{l-1}$

\emph{Uniform crossover} Schema $h$ survives iff none or all bits
swapped together. $p_{survival}=p^{o(h)}_x + (1-p_x)^{o(h)}$.

\emph{Distruption factor} $\epsilon(h,t)$ by recombination:
$\epsilon(h,t)=P_c\cdot(1-p_{survival})$

\subsubsection{Theorom}
Selection mutation and recombination combined:
$m(h,t+1)\leq m(h,t)\phi(h,t)[1-\epsilon(h,t)]$.
Net growth factor: $\gamma(h,t)=\frac{m(h,t+1)}{m(h,t)}$.

\subsection{Fitness Correlation coefficents}
$F_p$: mean fitness of the parents. $F_c$: mean fitness of the children.
$\sigma(F_p) =$ standard deviation of fitness parents
$\sigma(F_c) =$ std. dev. of children. coveriance fitness between parents and children:
$cov(F_p,F_c)=\sum^n_{i=1}\frac{(f(p_{gi})-F_p)(f(c_{gi})-f_c)}{n}$.
Operator fitness correlation coeficient
$\rho_{op}=\frac{cov(F_p,F_c)}{sigma(F_p)\sigma(F_c)}$.

\section{ES representation}
Fitness function: $f(x_1,\dots,x_n):\mathbb{R}^n\to \mathbb{R}$.
Genotype representation of an individual solution:
$(x_1,\dots,x_n,\sigma^2_1,\dots,\sigma^2_n, c_{12},\dots,c_{n-1,n})$,
paremeters $(x_1, \dots,x_n)$ need to be optimized.
Individiual solution consists of 3 parts: $\vec{x}$: problem vairables
$\Rightarrow$ Fitness $f(\vec{x})$. $\vec{sigma}$: standard deviations
$\Rightarrow$ variances. $\vec{\alpha}$: rotation angles $\Rightarrow$
covariances. The $n$-dimensional normal probability density function:
$p(X=x_1,\dots,x_n)=\frac{exp(-\frac{1}{2}X^TC^{-1}X)}{\sqrt{(2\pi)^n|C|}}$.
$C$: correlation matrix ($c_{ij}$; $|C|$ determinant $\Rightarrow$ rotation 
angles $\alpha_{ij}$: tan $2\alpha_{ij}=2c_{ij}/(\sigma^2_i-\sigma^2_j)$.
1-dimensional gaussian function:
$f(x)=\frac{1}{\sqrt{2\pi \sigma^2}}exp\frac{(x-\mu)^2}{2\sigma^2}$
\subsection{Mutation}
Complete covairance matrix: $\sigma_1,\dots,\sigma_n$; $c_{ij}\neq 0 (i\neq j)$.
The strategy paremeters are mutated:
$\sigma'_i=\sigma_ie^{
	\frac{\mathbb{N}(0,1)}{\sqrt{2n}}+
	\frac{\mathbb{N}_i(0,1)}{\sqrt{2\sqrt{n}}}
}$
$\alpha'_j=\alpha_j+\beta\mathbb{N}_j(0,1)$.
$\beta = 5$ degree in radians. $\mathbb{N}(0,1)$: standard normal
distribution. Then the problem paremeters are mutated with the new
strategy: $\vec{x}'=\vec{x}+\vec{N}(\vec{0},\vec{\sigma},\vec{\alpha}')$.
$\vec{N}$: n-dimensional normal distribution.
Often this is simplified by not doing angles or even just having a single
standard deviation (ie no index at $\sigma$ and nothing after the + 
in the power $e$).
\subsection{Recombination}
create one offspring from several parents selected randomly. Problem parameters
are averaged. The std devation is selected from one of the parents. Rotation
angles are not recombined
\subsection{Selection}
ES applies a high selection pressure: from $\mu$ parents $\lambda$ offspring
are generated with $\lambda >> \mu$ (~5 to 10 times). The best $\mu$ solutions
of the $\lambda$ offspring are selected, (or from both offspring and parents,
but experiments show just offspring has better selfadaptation).
\section{Ada pursuit}
Given: Set of $K$ \emph{operators} $A=\{a_1,\dots,a_k\}$. 
\emph{Probability vector} $P(t) = \{P_1(t),\dots,P_k(t)\}$. operator $a_i$ 
applied at time t in proportion to probability $P_i(t)$. env. returns rewards
$R_i(t) \geq 0$. Goal: Adapt $P(t)$ so that expected value of cumalitive
reward $\varepsilon[R]=\sum^T_{t=1}R_i(t)$ is maximized.

\subsection{Probabilty matching}
update $P(t)$ so probability of applying operator $a_i$ matches
the proportion of the estimated reward $Q_i(t)$ to the sum of all reward 
estimates $\sum^K_{a=1} Q_a(t)$
\subsubsection{PM: Reward estimate}
compute an estimate of the rewards, and older rewards should be less
influantial in non-stationary envs:
$Q_a(t+1)=Q_a(t)+\alpha[R_a(t)-Q_a(t)]$

\subsubsection{PM: Probability adaptation}
In non-stationary envs the probabilty of applying
any operator should never be less than some minmal \emph{thershold} $P_{min} > 0$.
For $K$ operators maximal probablity $P_{max} = 1-(K-1)P_{min}$. updating rule
for $P(t)$: $P_a(t+1)=P_{min}+(1-K\cdot P_{min})\frac{Q_a(t)}{\sum^K_{i=1}Q_i(t)}$.
\subsubsection{PM: Problem}
If one operator is consistently better probablity will still converge to average.

\subsection{Adaptive Pursuit strategy}
Update $P(t)$ so \emph{operator} $a^*$ that currently has the max estimated
reward $Q_{a^*}(t)$ is \emph{persued}. To achieve this the pursuit method
\emph{increases} selection probability $P_{a^*}$ and decreases all other
probabilities $P_a(t),\forall a\neq a^*$. So here it can deal with non-static envs.
\subsubsection{APS: Adaptive method}
Similiar to probability matching:
$Q_a(t+1)=Q_a(t)+\alpha[R_a(t)-Q_a(t)]$. Different from probablity matching:
Selection probability $P(t)$ is adapted in a greedy way.
\subsubsection{APS: Probability adoptation}
The selection probability of the current best operator
$a^*=argmax_a[Q_a(t+1)]$ is \emph{increased} $(0<\beta<1)$:
$P_{a^*}(t+1)=P_{a^*}(t)+\beta[P_{max}-P_{a^*}(t)]$.
The others are \emph{decreased}: 
$\forall a\neq a^*: P_a(t+1)=P_a(t)+\beta[P_{min}-P_a(t)]$

%TODO: the psuedo code??

\section{Convergence models}
Number of fitness function evaluations is proportional to the
number of generations run and pop size $FitFunc.count = Gen.count * Popsize$.
Selection algorithm and variation operators can be ignored.

\emph{Convergence speed:}
High selection presure is faster covnergence. Bigger popsize give better solutions.
Minimal fitness function evals is a tradeoff between selection presure and popsize.
\subsection{Selection Intensity}
Quantifies the speed of convergence. $I$ selection intensity. Response to
selection $R(t)$ is difference between mean fitness of generation $t+1$ and fitness
of generation $t$. Selection differential $S(t)$ is the diff between parent pop
at generation $t$ and mean fitness of generation $t$. 
$S(t)=\overline{f(t^s)}-\overline{f(t)}$. Assuming the pop fitness is normally
distributed $N(\overline(f),\sigma^2)$ we can scale the selection differential
by thestandard deiation $\sigma(t)$. This scaled selection differential is 
called selection intensity $I(t)$. This is a dimensionless number since 
the std deiation has the same units as the selection response:
$I(t)=\frac{S(t)}{\sigma(t)}=\frac{\overline{f(t^s)}-\overline{f(t)}}{\sigma(t)}$.
Standerdizing the normal distribution $(\overline(f)=0,\sigma=1)$ shows that
selection intensity $I$ is simply the epected average fitness
of the population after applying the selection scheme to a population
with standardized normal distibuted fitness $(N(0,1))$.
The relation between the response to selection $R$ and the selection
differential $S$ is given by the heritability $h^2$: $R(t)=h^2S(t)$ or
$R(t)=h^2\sigma(t)I(t)$.

\subsection{Proportionate selection}
$P_i(t)$ the proportion of occurences of individual $i$ at generation $t$.
$i$ has fitness $f_i$ and the mean fitness of generation $t$ is $\overline{f(t)}$.
$P_i(t^s)$ the proportion of individual $i$ in the parent pool
after applying proportionate selection: $P_(t^s)=P_i(t)\frac{f_i}{\overline{f(t)}}$.
The selection differential of proportiaonate selection is: 
$S(t)=\overline{f(t^s)}-\overline{f(t)}=\frac{\sigma^2(t)}{\overline{f(t)}}$
The selection intensity $I(t)=\frac{S(t)}{\sigma(t)}$ is tus eaual
to the ratio of the std devation of the fitness and the pop mean fitness:
$I(t)=\frac{\sigma(t)}{\overline{f(t)}}$

\end{multicols}
\end{document}