### 2016-01-30 evolutionaryComputation/2-practical/report.tex
adding more text and both tables

\documentclass{article}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage{booktabs}
\usepackage{color}
\usepackage{bussproofs}
\usepackage[pdftex]{graphicx}
\DeclareGraphicsRule{*}{mps}{*}{}
\usepackage{pifont}
\usepackage{qtree}
\usepackage{emp}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{svg}
\usepackage{pgfplots}
\usepackage{pgfplotstable}

\pgfplotsset{
    compat=1.9,
    compat/bar nodes=1.8
}

\begin{filecontents*}{dataTable.txt}
zero 	name 		firsty 		secondy		errbotstart		errbotend 		errtopstart			errtopend
0 		MLS           0         	1			0.1				0.2				0.1	               0.2
0 		ILS           1         	2			0.3				0.1             1.5               0.4
0 		GLS           2         	3.5			0.5				0.2             0.3               0.4
\end{filecontents*}

\newenvironment{scprooftree}[1]%
{\gdef\scalefactor{#1}\begin{center}\proofSkipAmount \leavevmode}%
{\scalebox{\scalefactor}{\DisplayProof}\proofSkipAmount \end{center} }

\newcommand{\brcell}[2][l]{%
	\begin{tabular}[#1]{@{}l@{}}#2\end{tabular}}
\begin{document}
\begin{empfile}
\begin{empcmds}
input metauml;
\end{empcmds}
\author{Jappie Klooster}
\title{Graph bipartitioning and you}
\maketitle

\section{Introduction}
This is a report of the evolutionary practical assignment 2.
In here I will discuss the execution of the assignment, what it was, which
methodology was used to execute the assignment.

The intended audience is people familiar with programming.

\section{Assignment}
Graphbipartitioning is dividing a graph into 2, with as little edges between
both graphs as possible.

\section{Methodology}
I coppied all the code from the previous practical into this new folder.
Then I went onto a delete rage. Most code was deleted except the Expirment
class, this will probably still be usefull.

I use git, so it doens't really matter to accidently delete to much. After
that I changed the data, I removed everything which was useless. Now the file
only contains the connection numbers. This makes parsing the file a lot
easier. I don't want to do advanced pattern matching when I can quickly modfy
the file with a few vim macros. (in this case I just did a `30x').

\subsection{MLS}
MLS basically involved implementing the initial vertex swap first improvement
local search and tying it into the existing framework. I did this by just
using the existing createPopMethods with a new solution generator. Then
I just mapped this population with the local search (so the type is
$IMember \to IMember$, but the solution changes).

\subsection{ILS}
While implementing ILS, I had it turn out results much worse than MLS.
The initial implmentation would just stop if it found a worse local optimum.
The description sais it should go back to the previous optimum. So I guess
that you should try again after that? But then I wonder how often you should
try again.

I got much better results with a system where the local search got several chances.
If it failed it would fallback, and then try again. 
With chances around 30 I managed to get up and around 1900.
When I reset the chance counter on a succesfull find
I managed to get consistently 2200, one result was as high as 2310.

I don't think there is a way to do it any better without drastically making
the algorithm more complex.
This current search for improvement was just based on the fact that ILS was
underperforming to MLS, which shouldn't be the case according to my memory
from class. Now its overperforming quite drastically so I'm happy.

\subsection{GLS}
I implemented GLS with the evolution class. The only thing I had to do really
was write the crossover method, the rest was already functional.

I wanted to see if the GLS popsize and selection pressure would have any
effect on the average fitness found. I suspect that a bigger population size
would find better results and that weaker selection would find better
average scores.

Here I also started to think about my own topic to investigate, at first
I just wanted to do varying population sizes and varying selection methods.
But I thinks this takes a little to much time and it doesn't quite fit
on the tTest table so I decided the varying popsizes (spoilers: it produces
better solution at the cost of exponentially more time).

I chose to investigate with population sizes of 25, 50 and 75. I didn't want
to wait to long on this and the runtime for 50 was approaching ~10 minutes
for the slowest. Luckily I have a powerfull laptop and can use paralization.

Doing this 30 times to cancel the staostic effect would result into ~5 hours.

\subsection{Counting double}
I had an issue in which my classmates would get much better results than me.
After discussing it for a while it turned out that I was counting double. However
This didn't make sense because I just used scala combinators to count the 
elements. Basically this means that there would have to be a bug in the std
Scala library which would seem highly unlikely. I knew I was counting double
because all my results where even numbers. After thinking about it for a while
I decided to investigate the graph file, it turns out that every connection
is made double, so the double counting was in fact correct.

\subsubsection{Graph pruning}
I decided to use a prune graph to speed up the algortihm. In the beginning I
create 2 graphs. One normal and one pruned, The pruning works by selecting
comparing all connections against the id, if lower we ignore the connection.
To make sure we only do this on doubly linked graphs 2 graphs are created.
If the prune graph has exactly half the score of a random solution we assume
its a doubly linked graph so we can prune it.
The pruning effectivally halfs the ammount of connections that need to be
checked by each itteration (this happens in the evaluation). The
FidduciaMathesis has even more gain because it has a lot of book keeping to
do with the connections (although it tries to reduce this with the freecell
array, having a smaller initial graph is even better).

\subsection{Fidducia Mathesis}
I was somewhat sceptical of my local search implementation. I had quite 
different results from my classmates so to compare I decided to implement
the fidducia mathesis local search to. It was quite a lot of work but it
truly is significantly faster.

Finding the paper was somewhat difficult as the original was behind a paywal.
But somoene put a copy on github so I just used that one. The algorithm 
involves a lot of cleaver book keeping:

\section{TODO}
MLS with several results
GLS -> I think the localsearch will recursively unecesrly decend into its members
because the local search is packed in the member factory. Maybe remove the
local search from the memberfactory?

\section{Results}
\subsection{Fiduccia Mathesis}
\subsubsection{Time}
\pgfplotstableread{fidRuntimes.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={fidRuntimes.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Time}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};

\end{axis}

\end{tikzpicture}

\input{fidRuntimes.textable}

\subsubsection{Optima}
\pgfplotstableread{fidResults.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={fidResults.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Fitness (less is better)}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};

\end{axis}

\end{tikzpicture}
\\
\input{fidResults.textable}

\subsection{Vertex swap first improvemnt}
\subsubsection{Time}
\pgfplotstableread{vertRuntimes.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={vertRuntimes.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Time}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};

\end{axis}

\end{tikzpicture}

\input{vertRuntimes.textable}

\subsubsection{Optima}
\pgfplotstableread{vertResults.table}\dataTable
\begin{tikzpicture}
\pgfplotsset{set layers}
\begin{axis}[
	ybar stacked,
	xlabel={Algorithm},
	xticklabels from table={vertResults.table}{name},
	xtick=data,
	x tick label style={rotate=90,anchor=east},
	ylabel={Fitness (less is better)}
]
% Layers needed to get error bars on top of lower blue rectangle
\addplot[color=white]
table[on layer=axis grid, x expr=\coordindex, y=firsty]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=zero, y error minus=errbotstart, y error plus=errbotend]{\dataTable};

\addplot[color=red] plot[on layer=axis lines,error bars/.cd, y dir=both, y explicit] 
table[x expr=\coordindex, y=secondy, y error minus=errtopstart, y error plus=errtopend]{\dataTable};

\end{axis}

\end{tikzpicture}
\\
\input{vertResults.textable}
\end{empfile}
\end{document}