### 2016-04-25 multiAgentLearning/summeries/InteractiveLearningProblem.tex
new subject!!

\documentclass{article}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\algrenewcomment[1]{\(\triangleright\) #1}
\begin{document}
\author{Jappie Klooster}
\title{A summery of the interactive learning problem}

\maketitle

Equilibrium is very important to social systems. The nearest
thing we have as a fully developed framework for the study of
equilibria in social systems is Bayesian decision theory.

However there are difficulties in using this. Firstly it requires that
everyone has to anticipate correctly the behaviours of others.
Secondly it requires that individuals improve their strategies
at each point in time as events unfold.
Both these things are incredibility difficult if not impossible to do.

The reason why this only affects social systems and not physical systems
(which also desire equilibria) is that they're self referential.
The act of learning changes the thing to be learned.

\end{document}