### 2016-04-26 multiAgentLearning/summeries/learningAndTeaching.tex
summeries

\documentclass{article}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\algrenewcomment[1]{\(\triangleright\) #1}
\begin{document}
\author{Jappie Klooster}
\title{A summery of the interactive learning problem}

\maketitle

A lot of AI involves learning performed by a single agent. In this case the
agent tries to figure out a function. Many techniques have been
developed for this.

However in a multi-agent system the environment contains other
agents that also try to learn a function. Thus the environment
changes without the agent interacting with it. So the learned
function may be wrong.

Simple rules can lead to complex behavior, this leads us to believe
that one cannot separate learning from teaching in multi agent systems.
When deciding what to do, an agent should consider both the things he
learned from other agents passed behavior and how he wants to
influence their future behavior.

A problem arises when both agents want to be play teacher and
both keep selecting uncoordinated actions. The question is if
we can think of a simple rule that will enable them
to coordinate without an external designation of a teacher.

We consider something learning if there is a temporal nature in the domain
and regularity across time. By doing this we can used
passed experience to improve future strategy.

Tit for That is an example of a very simple learning strategy.
You could think of more complex strategies in which you consider
more of the history to play in a more sophisticated manner.
Consider frequency analysis of the opponents play for example.
This is called fictitious play.

Stochastic games are another example of where learning can be
applied.

Other aspects involving both types are weather agents already
know the game. If the game is not known the agents can in
addition learn about the structure of the game.
Another question is if the game is observable, do the
players know each others actions/payoff while playing?

Another model that won't be treated is the investigation
of large populations, this is used in biology.

Learning strategies should be understood as choosing actions
and updating beliefs. A consequence however is that
using `accumulated knowledge' may not always be beneficial
because it pre commits to a given strategy (while others may
be better).

Descriptive theories try to tell us how learning takes place
in real life.
The ideal descriptive theory would have the following two properties:
Realism: `There should be a good match between the formal theory and the
natural phenomenon being studied.', Convergence: `The formal
theory should exhibit interesting behavioral properties, in particular
convergence of the strategy profile being played to some solution concept
(e.g., equilibrium) of the game being played.'.
To demonstrate realism we can use experiments.

The holy grail of convergence properties is demonstrating convergence to
stationary strategies which form a Nash equilibrium of the stage game.
Another approach recognizes Nash equilibria are rare as a result
of a learning process. It requires the empirical frequency of
play convergence to such an equilibrium.
The third approach is giving up on Nash equilibria. This uses
convergence to a correlated equilibrium of the stage game.
The last one gives up on convergence of stationary policies. Instead it
requires that the non-stationary policies converge to an interesting state.

Prescriptive theories describes how learning should happen.
For zero-sum games and even in repeated or stochastic
zero sum games it can be asked if learning is happening optimally.
But in general this won't work because the learning often depends on the
learning of other agents.

The equilibrium of learning strategies is an uncommon approach
instead we use the more common and more practical approach
of asking weather the payoff of a learning strategy is good enough.

There are multiple different payoff requirements:
Safety: `A learning rule is safe if it guarantees the agent at least
its maxmin payoff, or “security value.” (Recall that this is the payoff the agent
can guarantee to himself regardless of the strategies adopted by the opponents)'
, Rationality: `A learning rule is rational if whenever the opponent
settles on a stationary strategy of the stage game (i.e., the opponent adopts the
same mixed strategy each time, regardless of the past), the agent settles on a best
response to that strategy.', No-regret: `Against any set of opponents it yields a payoff that is no less than the
payoff the agent could have obtained by playing any one of his pure strategies
throughout. We give a more formal definition of this condition later in the chapter.'
\end{document}