### 2016-04-26 multiAgentLearning/notes.tex
notes...

\documentclass{article}
\usepackage{algorithmicx}
\usepackage{algpseudocode}
\usepackage{enumerate}
\renewcommand{\thesubsection}{\thesection.\alph{subsection}}
\algrenewcomment[1]{\(\triangleright\) #1}
\begin{document}
\author{Jappie Klooster}
\title{A note on multi agent learning}

\maketitle

Almost surely or a.s for short \} $pr(\dots) = 1$, but this $\neq$ certainity,
because you cannot proof it.

11111111
1211111
41343434
5311111

DIagonal argument. You can use the diagonal to realize
that .... .... .... bla bla not equal subset is equal to 0??

if you can read everything and there is no possiblity of a tie than in a
voting game there is convergence in one round.

Peyton young does some qualititatively analyses.
\end{document}