### 2016-06-29 multiAgentLearning/paper/article.tex
added a fuckton of text

\abstract{
  %TODO abstract
  \todo[inline]{abstract missing}
}
\todo{be more critical? maybe it depends to much on its paremeters?
  Or which game types are good for saticifing play?
}
\todo{misses callouts, intermezzi and quotes}
\section{Introduction}
\label{sec:introduction}
\toReview
Satisficing play was originally introduced
to try and limit the computation power that earlier models required of human
behavior \citep{herbert}. Originating from the field of economics, they
wanted to use a model that was not ``super intelligent''.

\toReview The classical models from the field of economics had at least one of the
following assumptions:
\begin{itemize}
\itemsep0em 
  \item The structure of the game is known.
  \item Awareness of other player's decisions.
  \item Awareness of being in a game situation
  \item Being able to preform maximization calculations.
\end{itemize}

\toReview These assumptions can be called common knowledge, and are a demanding to
make \citep{binmore}.

\todo{give paper overview}

\section{Core algorithm}
\label{sec:core-algorithm}
\toReview The core algorithm as presented by \citep{herbert} and formalized by
\citep{karandikar},
in here we present the simplified version of \citep{stimpson:2001} first:

\begin{enumerate}
  \item $ \begin{cases}
              \text{if } \pi_t \geq \alpha_t & A_{t+1} = A_t\\
              \text{else} & A_{t+1} \neq A_t
              \end{cases}$
  \item $ \alpha_{t+1} = \lambda \alpha_t + (1-\lambda) \pi_t $ \\
\end{enumerate}
\todo{add table caption}
\begin{tabular}{ll}
  Definition & symbol \\ \toprule
  Time & $t$ \\
  Action & $A_t$ \\
  Utility & $\pi_t$ \\
  Aspiration & $\alpha_t$ \\
  Learning rate & $\lambda$ \\ \bottomrule
\end{tabular}

\toReview In the first phase the next action is selected. If the action still holds
according to the satisficing level or ambition it will remain the same, if not another
action will be selected. In the second phase the satisficing level is updated.
In here the constant learning rate is used. This is a variable that can be
modified to adjust the amount of new information being used or how much
information from memory is used.

\section{Satisficing play in action}
\todo{make an intermezzo of this?}
\toReview
To demonstrate a potential use of this algorithm we will use at several example
cases. The first one being the iterated prisoner's dilemma.
\fbox{
  \begin{minipage}{2.5in}
    \toReview
    \subsection{A prisoner's dillema introduction}
    The prisoner's dillema is one of the classical game theory examples.
    Consider the following normal form game which is a prisoner's dillema:
    \begin{tabular}{lll}
                & Cooperate   & Defect \\
      Cooperate & $(2,2)$     & $(0,3)$  \\
      Defect    & $(3,0)$     & $(1,1)$  \\
    \end{tabular}
    The Nash equilibrium of this game lies on (Defect,Defect). However, all but
    that action are pareto optimal actions. So the Nash equilibrium would be the
    ``right'' strategy to play if the game is only played once. However if you repeat
    the game against the same opponent you can ``solve''
    the prisoner's dilemma, meaning not playing the Nash equilibrium,
    but trying to get pareto optimal rewards. This
    is often done by starting defecting as soon as your opponent starts
    defecting, both the tit for tat and win stay lose shift strategies do this.
  \end{minipage}
}
\toReview Strategies such as tit for tat or win stay lose shift already know
which actions are winning and which are punishing to try and solve it.
Satisficing play learns this however by updating the aspiration it got from
playing a move. If we consider a game versus always cooperate for example:

\begin{tabular}{lllll}
    t & Cooperate & $A_t$ & $\pi_t$ & $\alpha_t$ \\
    0 & C     & C    & 2       & 3 \\
    1 & C     & D    & 3       & 2.5 \\
    2 & C     & D    & 3       & 2.75 \\
\end{tabular}

\toReview In $t=0$ satisficing play doesn't meet its ambition of 4, therefore, in the
next game it modifies its ambition and changes its action. Here we can see
that satisficing play does get more utility than its ambition, and is satisfied
with its action.
But now if we start with a lower apsiration level:

\begin{tabular}{lllll}
  t & Coop & $A_t$ & $\pi_t$ & $\alpha_t$ \\
  0 & C     & C    & 2       & 1 \\
  1 & C     & C    & 2       & 1.5 \\
  2 & C     & C    & 2       & 1.75 \\
\end{tabular}

\toReview So here satisficing play finds a Pareto optimal action, but it doesn't find
the best possible action it could get against this opponent. From this example
we can also see how dependent satisficing play is on its starting parameters.

\toReview Now consider a game versus tit for tat (TT), Tit for Tat is a
strategy that always copies the last action of its opponent.
This game can also be found in the \citep{stimpson:2001} paper:

\begin{tabular}{lllll}
        t & TT & $A_t$ & $\pi_t$ & $\alpha_t$ \\
        0 & C     & C    & 2       & 3  \\
        1 & C     & D    & 3       & 2.5 \\
        2 & D     & D    & 1       & 2.75 \\
        3 & D     & C    & 0       & 1.87 \\
        4 & C     & D    & 3       & 2.435 \\
        5 & D     & D    & 1       & 1.715 \\
        6 & D     & C    & 0       & 0.85875 \\
\end{tabular}

In this table we can see satisficing play and tit for tat ending up in a cycle.
at $t=0$ satisficing gained to little reward for its ambition level $\alpha_t$,
and therefore switches action in the next round. In $t=1$ saticificing play
is satisfied, but int $t=2$ it no longer is so it switches but this even ends
up being worse at $t=3$. After that the cycle resets at $t=4$, the cycle becomes
most obvious when looking at the $\pi_t$ column. Also note about this opponent
that if saticificing play would've started with a lower aspiration, it would
not enter this cycle and just keep on cooperating.
\fbox{
  \begin{minipage}{2.5in}
    \toReview
    \subsection{Karandikar's algorithm}
    In the \nameref{sec:core-algorithm} section we said we would use Karandikar's
    notation, but use the simplified version of stimpson.
    Karandikar version would not necessarily change action when being
    disappointed (aspiration level not met), but would switch with a
    probability $p$, which he called \emph{inertia}.

    He also introduced another model, which used random perturbations on
    $\alpha_t$. This ``solved'', the initial state problem, however to prevent
    the random trembles to run out of hand relative to the game rewards the tremble
    generating function needed a feasible payoff interval $\Lambda$. Which
    requires more knowledge of the game, breaking the initial assumptions
    presented in the \nameref{sec:introduction}.
  \end{minipage}
}

\section{A saticificing comparison}
\drafting So, until now we've seen thoroughly how saticificing play works, and even some
strengths and weaknesses. However, what we really want to see is how it does
compare against other MAL algorithms. So to do this we'll first introduce the
competitors (Q-learning and WOLF-PHC) and then we'll show per game how they
do in comparison with help of the work of \citep{crandall}.

Q-Learning is described by \citep{sandholm} in the following way: ``The
Q-learning algorithm works by estimating the values of state-action pairs. The
value $Q(s,q)$ is the expected discounted sum of future payoffs obtained by taking
action $a$ from state $s$ and following an optimal policy thereafter.
Once these values have been learned, the optimal action from any state
is the one with the highest Q-value. After initializing the Q-values to arbitrary
numbers they are estimated based on experience as follows: 1. Select an action
$a$ from the current state $s$, causing an immediate payoff $r$ and arriving at
state $s'$. 2. Update $Q(s,a)$ based on this experience as follows:
\[\Delta Q(s,a) = \lambda[r+\gamma \text{ max }Q(s',b)-Q(s,a)]\]
Where $\lambda$ is the learning rate and $0 \leq \gamma < 1$ is the discount
factor.
3. Go to 1.
'' A look-up table is used to store the Q-values.

\todo{explain WOLF-PHC better?}
WoLF-PHC stands or win or learn fast policy hill climbing. It is 
introduced as an extension to q-learning. However with the addition that
learning rate should be increased when ``losing''. A more precise definition
can be found in the paper of \citep{bowling}.

The actual comparison of the algorithms is done by \citep{crandall}, in here
we will discuss several of his results. All the algorithms are compared in self
play. The first game result we'll discuss
is the already introduced prisoner's dillema:
\includegraphics[width=2.5in]{crandall/prisoners}

Where the S-Algorithm is saticificing play as defined by stimpson, WoLF-PHC is
defined as above and QL is Q-learning. We can clearly see saticificing play
performing a lot better than the other two algorithms in this instance.
(Note that in this figure different values were used than our example, but
the basic structure of the game is the same).

A game of chicken is a game where 2 players drive in a straight line in their
cars straight towards each other. If both players keep on driving straight,
they'll crash and get a bad reward. If one of them moves out of the way, he'll
be ashamed, but be better of than crashing, the other one will of course get
some fame, and if both players move out of the way, they'll share the same.
Formally it looks like this:

\begin{tabular}{lll}
              & Bend & Straight \\
    Bend & $(3,3)$     & $(2,3.5)$  \\
    Straight & $(3.5,2)$     & $(1,1)$  \\
\end{tabular}

And here is the result of Crandall:
\includegraphics[width=2.5in]{crandall/chicken}

In here we see again saticificing play performing better than Q-Learning and
WoLF. However, the difference is a lot less refined.

Shapley's game is a 3 action choice game. If saticificing play has to change
action, instead of picking the other action, it will just pick uniformly random
of the 2 other actions. The context of the game is left to the imagination
of the reader and the normal form game looks like this:

\begin{tabular}{llll}
              & a & b & c \\
    A & $(0,0)$ & $(0,1)$     & $(1.0)$  \\
    B & $(1,0)$ & $(0,0)$     & $(0.1)$  \\
    C & $(0,1)$ & $(1,0)$     & $(0.0)$  \\
\end{tabular}

This is the result of crandall:
\includegraphics[width=2.5in]{crandall/shapleys}

In here we can see saticifing play performing a lot worse than the other two
algorithms. This is because, Crandall explains, the joint aspiration never
falls below the Pareto frontier, thus the behavior is essentially random.

\section{Properties of satisficing play}
\drafting

\todo{
Using both \citep{karandikar} and \citep{stimpson:2001} works to look deeper
at the properties of satisficing play.

karandikar used a theoretical proofing method, whereas Stimpson showed some
experimental results.

also showing the contrast graph which supports the work of \citep{arthur}
}
\section{An extended satisficing play}
\drafting

\todo{
In here we discuss the work of \citep{stimpson:2003}, where he extends the
algorithm to be played with mutliple action choices.
Stimpson created a game where resources had to be allocated among agents.
}

\todo{add caption for tables?}
\bibliography{main}

\address{Jappie Klooster\\
  Dept. of Informatics\\
  Universiteit Utrecht\\
  The Netherlands\\}
\email{j.t.klooster@students.uu.nl}

%  LocalWords:  Karandikar